GRF CF Loop

### 2025-11-17 Writing/manuscript_CF_continuous.qmd
comments on the manuscript added

---
title: "Estimation of the effect of a continuous treatment variable"
format:
  docx:
    toc: true
    number-sections: true
    number-tables: true 
  html:
    toc: true
execute:
  echo: false
bibliography: references.bib
---

@@

Efficient nitrogen (N) management remains one of the most critical challenges in modern agricultural production. Nitrogen is essential for achieving high crop yields, yet excessive or poorly timed applications can lead to economic losses and environmental harm. The principle of applying the right amount of nitrogen at the right place and time requires understanding the spatial and temporal variability in crop nitrogen needs. This variability arises from complex interactions among soil nitrogen availability, soil and field characteristics, and weather conditions. Accurately identifying the optimal nitrogen rate for each site and season is therefore fundamental to improving both profitability and sustainability in crop production systems.

On-farm field experiments have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management. Through on-farm agronomic trials, researchers can directly observe yield responses to varying nitrogen rates under real-world field conditions. Analyzing these data enables estimation of yield response functions and identification of site-specific economically optimal nitrogen rates. Recent advances in data analytics—particularly the widespread adoption of machine learning (ML) methods—have expanded the potential of on-farm experimentation. Two major methodological directions have emerged: (1) yield prediction–based methods, which recommend site-specific input rates by predicting yield as a function of nitrogen and environmental covariates (@krause2020random; @barbosa2020modeling; @barbosa2020risk; @gardner2021economic), and (2) causal machine learning (causal ML) methods (@kakimoto2022causal), which estimate the underlying causal relationship between nitrogen application and yield response.

In the OFPE setting, causal machine learning methods have shown to be promising to recover the full yield response curve. @kakimoto2022causal showed that causal forest can estimate heterogeneous treatment effects of an input on yield more accurately than prediction-oriented machined approaches like random forest, boosted forest, and convolutional neural network. However, their application of CF is only limited to a single-year OFPE experiment from a single field. Causal forest was developped primarily to estimate
the heterogeneous treatment effects of a dicrete treatment. In OFPE, there are typically five
or six distinctive target experimental rates (see the first panel of @fig-distribution-N). In @kakimoto2022causal, CF estimates the effect of multiple treatments defined as the increase in nitrogen rate from the lowest rate to the other rates (so n − 1 tretments in n-rate trials). However,
once data from multiple fields with different target rates are combined, applied nitrogen rates
become rather continuous as seen in the second panel of @fig-distribution-N. Existing approaches—such as direct applications of random forests or neural networks and CF—often tend to fail to capture heterogeneous treatment effects of a continuous variable (@kakimoto2022causal).

@@

```{r fig-distribution-N, fig.cap="Distribution of nitrogen rates from a single OFPE and multiple OFPEs"}
knitr::include_graphics(here::here("writing/Figures/g_N_hist_single_multiple.pdf"))
```
As multi-year, multi-field OFPE datasets accumulate, pooling data across fields has become both feasible and desirable—it can improve estimation of yield response functions and deliver stronger agronomic and economic insights. It creates a need for methods that handle continuous treatments and it becomes especially important to adapt causal methods to continuous treatments without losing their statistical strengths over other machine-learning approaches.

@@

@@

@@

```{=tex}
\begin{align*}
Y=\theta(X) T+g(X, W)+v
\end{align*}
```

where $Y$ is the outcome, $T \in\{0,1\}$  is the treatment, $X$ covariates that also modify the treatment effect $\theta(X)$, $W$ additional predictors of $Y$, and $v$ is noise. 
Orthogonalization estimates $f(X, W)=\mathbb{E}[Y \mid X, W]$ and $h(X, W)=\mathbb{E}[T \mid X, W]$, forms residuals $Y_{\mathrm{res}}=Y-f(X,W)$ and $T_{\mathrm{res}}=T-h(X, W)$, and then relates $Y_{\mathrm{res}}$ to $T_{\mathrm{res}}$. Because this score is Neyman-orthogonal and is typically implemented with cross-fitting, small errors in $f$ and $h$ do not induce first-order bias in $\theta(X)$.

@@

```{=tex}
\begin{align*}
\mathbb{E}[Y \mid X, W]=\theta(X) \mathbb{E}[T \mid X, W]+g(X, W)=\theta(X) h(X, W)+g(X, W)
\end{align*}
```

@@

```{=tex}
\begin{align*}
Y_{\mathrm{res}}=Y-\mathbb{E}[Y \mid X, W]=\theta(X)(T-h(X, W))+v=\theta(X) T_{\mathrm{res}}+v
\end{align*}
```

@@

@kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield and treated heterogeneity in a discrete way. However, when data are pooled across fields and applied nitrogen rates become continuous, the standard residualization of the treatment in DML does not extend cleanly to that setting.

@@

```{=tex}
\begin{align*}
Y=\theta(X, T)+g(X, W)+v
\end{align*}
```

@@

```{=tex}
\begin{align*}
\theta(X, T)-\mathbb{E}[\theta(X, T) \mid X, W] \neq \theta(X, T-\mathbb{E}[T \mid X, W])
\end{align*}
```

and the clean decomposition that underlies double-orthogonalization in the linear-in-$T$ case breaks down. 
Instead, we apply single-orthogonalization (SO). We first estimate $f(X, W)=\mathbb{E}[Y \mid X, W]$ and form the residual $Y_{\mathrm{res}}=Y-f(X, W)$, and then recover $\theta(X, T)$ by flexibly modeling  $Y_{\mathrm{res}}$ as a function of $(X,T)$. This approach removes baseline variation due to $g(X,W)$ while preserving the full interaction structure between $X$ and the continuous treatment.

@@

```{=tex}
\begin{align*}
y = f(X, N) + g(W) + v
\end{align*}
```

@@

```{=tex}
\begin{align*}
\tilde{y}=y-\widehat{g}(W)
\end{align*}
```

@@

```{=tex}
\begin{align*}
\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
\end{align*}
```
where $P_{\text {Corn }}$ is the corn price, $P_{\text {N }}$ is the nitrogen price, and $\hat{f}$ is the estimated yield response.

@@

@@

```{=tex}
\begin{align*}
y=\sum_{k=1}^K \theta_k(X) S_k(N)+g(X, W)+v
\end{align*}
```

@@

```{=tex}
\begin{align*}
\tilde{y}=y-\mathbb{E}[y \mid X, W]=\sum_{k=1}^K \theta_k(X) \widetilde{S}_k(N)+v,
\end{align*}
```

@@

```{=tex}
\tilde{y}=y-\widehat{\mathbb{E}}[y \mid X, W], \quad \tilde{N}=N-\widehat{\mathbb{E}}[N \mid X, W]
\end{align*}
```

@@

```{=tex}
\hat{\tilde{y}}=\hat{\theta}(X) \cdot \tilde{N}
\end{align*}
```

@@

```{=tex}
\begin{align*}
RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
\end{align*}
```

@@

@@

```{=latex}
\begin{align} 
y_{i,j} = 
  \begin{cases}
  \alpha_{i,j} + \beta_{i,j} N + \gamma_{i,j} N^2 + \varepsilon_{i,j}, & N < \tau_{i,j} \\
  \eta_{i,j} + \varepsilon_{i,j}, & N \geq \tau_{i,j}
  \end{cases}
\label{eq:yield-gen-sim}
\end{align}
```

In this formulation, yield increases as $N_{i,j}$ increase until $N_{i,j}$ reaches $\tau_{i,j}$, at which yield hits the plateau, $\eta_{i,j}$. The yield level at $N_{i,j}= 0$ is $\alpha_{i,j}$. The rate at which yield increases with respect to $N_{i,j}$ is governed by $\beta_{i,j}$ and $\gamma_{i,j}$. For a given cell, the quadratic-plateu function has five parameters, but the degree of freedom is only three. That is, once any three of the five parameters are determined, then the rest of the parameters are automatically determined. Specifically, we first generated $\alpha_{i,j}$, $\eta_{i,j}$, and $\tau_{i,j}$ in a spatially correlated manner following the Gaussian bluh bluh process (cite), and then $\beta_{i,j}$ and $\gamma_{i,j}$ were derived. The erorr term ($\varepsilon_{i,t}$) is generated using the same spatial process. This is repeated independenty 500 times to create 500 fields of the same size but different spatial distribution and values of the three parameters.

@@

```{r fig-trial-design, fig.cap = "Example trial design"}
readRDS(here::here("writing/Figures/g_td.rds"))
```

@@

By contrast, NO_ANN exhibits substantial instability in the single-field case, with a heavy right tail and much higher errors, but rapidly converges toward the leading methods by 5–10 fields (mean RMSE improving to around 14–13), highlighting its strong dependence on sample size. SO_RF improves modestly with more fields (mid-20s RMSE for 10–20 fields) but remains consistently less accurate than SO_ANN or CF. NO_RF maintains broad, right-shifted distributions and higher mean errors throughout (≈ 36–40), confirming its role as a weak baseline. Finally, DO_ANN performs poorly throughout, with both histograms and means showing persistently high errors (≈ 90+) and negligible benefit from additional data, reflecting over-correction in the orthogonalization process.

@@

@@

# ggsave(here::here("results","rmse_mean_by_model_fields.png"), p_mean, width = 9, height = 5, dpi = 300)
```

@@

@@

best_models %>%
  gt() %>%
  tab_header(
    title = "Best models by field count",
    subtitle = "Ordered by lowest mean RMSE"
  ) %>%
  fmt_number(columns = c(mean), decimals = 3) %>%
  cols_label(
    field_f = "# fields",
    mean = "mean RMSE"
  ) %>%
  tab_options(table.font.size = "small")

@@

>>>>

---
title: "Estimation of the Effect of a Continuous Treatment Variable"
author:
  - name: Mona Mousavi
    orcid: 
    email: mmousavi2@huskers.unl.edu
    affiliations:
      - name: University of Nebraska Lincoln
  - name: Taro Mieno
    orcid: 0000-0002-0614-0771 
    email: tmieno2@unl.edu
    affiliations:
      - name: University of Nebraska Lincoln
  - name: Takashi Tanaka
    orcid: 0000-0001-7116-6962
    email: takashi@agro.au.dk
    affiliations:
      - name: Aarhus University
  - name: David Bullock
    orcid: 
    email: dsbulloc@illinois.edu
    affiliations:
      - name: University of Illinois at Urbana Champaign
format:
  html:
    toc: true
    number-sections: true
    number-depth: 3
  pdf:
    toc: true
    number-sections: true
    pdf-engine: xelatex
  docx:
    toc: true
execute:
  echo: false
bibliography: references.bib
---

@@

Efficient nitrogen (N) management remains one of the most critical challenges in modern agricultural production. Nitrogen is essential for achieving high crop yields, yet excessive or poorly timed applications can lead to economic losses and environmental harm (<span style = "color: blue;"> add citations here </span>). The principle of applying the right amount of nitrogen at the right place and time requires understanding the spatial and temporal variability in crop nitrogen needs. This variability arises from complex interactions among soil nitrogen availability, soil and field characteristics, and weather conditions. Accurately identifying the optimal nitrogen rate for each site and season is therefore fundamental to improving both profitability and sustainability in crop production systems.

On-farm field experiments have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management (<span style = "color: blue;"> add citations here. e.g., Dave's paper in agronomy and Lacoste's nature paper </span>). Through on-farm agronomic trials, researchers can directly observe yield responses to varying nitrogen rates under real-world field conditions. Analyzing these data enables estimation of yield response functions and identification of site-specific economically optimal nitrogen rates. Recent advances in data analytics—particularly the widespread adoption of machine learning (ML) methods—have expanded the potential of on-farm experimentation. Two major methodological directions have emerged: (1) yield prediction–based methods, which recommend site-specific input rates by predicting yield as a function of nitrogen and environmental covariates (@krause2020random; @barbosa2020modeling; @barbosa2020risk; @gardner2021economic), and (2) causal machine learning (causal ML) methods (@kakimoto2022causal), which estimate the underlying causal relationship between nitrogen application and yield response.

::: {.column-margin}
Right now, it is written as if prediction-oriented and causal ML have been used in parallel. But, @kakimoto2022causal is in response to the danger of applying simply applying prediction-oriented ML approaches. Modify the text to reflect that here. Cite [this](https://link.springer.com/article/10.1007/s11119-024-10136-x) to motivate the use of causal ML.
:::
<!--end of column-margin-->

In the OFPE setting, causal machine learning methods have shown to be promising to recover the full yield response curve. @kakimoto2022causal showed that causal forest can estimate heterogeneous treatment effects of an input on yield more accurately than prediction-oriented machined approaches like random forest, boosted forest, and convolutional neural network. However, their application of traditional CF is only limited to a single-year OFPE experiment from a single field because it was developed primarily to estimate
the heterogeneous treatment effects of a discrete treatment. In OFPE, there are typically five or six distinctive target experimental rates (see the first panel of @fig-distribution-N). In @kakimoto2022causal, CF estimates the effect of multiple treatments defined as the increase in nitrogen rate from the lowest rate to the other rates (so n − 1 treatments in n-rate trials). However, once data from multiple fields with different target rates are combined, applied nitrogen rates become rather continuous as seen in the second panel of @fig-distribution-N. While existing prediction-oriented ML approaches—such as direct applications of random forests or neural networks — can naturally handle continuous treatment variable, they tend to fail to capture heterogeneous treatment effects of a treat variable (@kakimoto2022causal). As multi-year, multi-field OFPE datasets accumulate, pooling data across fields has become both feasible and desirable—it can improve estimation of yield response functions and deliver stronger agronomic and economic insights. It creates a need for methods that can estimate heterogeneous treatment effects of a continuous treatment variable.

@@

```{r fig-distribution-N, fig.cap="Distribution of nitrogen rates from a single OFPE and multiple OFPEs"}
knitr::include_graphics(here::here("writing/Figures/g_N_hist_single_multiple.pdf"))
```

@@

::: {.column-margin}
<span style = "color: blue;"> We need to emphasize that we are proposing approaches that have not been tried anywhere: SO-NN and semi-parametric-CF. They are the main approaches. We then have other contrasting approaches to compare against. </span>
:::
<!--end of column-margin-->

@@

::: {.column-margin}
<span style = "color: blue;"> SO-ANN not introduced yet. You need to talk more about what you do in the above paragraph (mostly SO-NN and semi-parametric CF). </span>
:::
<!--end of column-margin-->

@@

$$
Y=\theta(X) T+g(X, W)+v
$$

where $Y$ is the outcome, $T \in\{0,1\}$  is the treatment, $X$ covariates that also modify the treatment effect $\theta(X)$, $W$ additional predictors of $Y$, and $v$ is noise. 
Orthogonalization estimates $f(X, W)=\mathbb{E}[Y \mid X, W]$ and $h(X, W)=\mathbb{E}[T \mid X, W]$, forms residuals $Y_{\mathrm{res}}=Y-f(X,W)$ and $T_{\mathrm{res}}=T-h(X, W)$, and then relates $Y_{\mathrm{res}}$ to $T_{\mathrm{res}}$. Because this score is Neyman-orthogonal and is typically implemented with cross-fitting, small errors in $f$ and $h$ do not induce first-order bias in $\theta(X)$ [(@chernozhukov2018double].

@@

$$
\begin{align*}
\mathbb{E}[Y \mid X, W]=\theta(X) \mathbb{E}[T \mid X, W]+g(X, W)=\theta(X) h(X, W)+g(X, W)
\end{align*}
$$

@@

$$
\begin{align*}
Y_{\mathrm{res}}=Y-\mathbb{E}[Y \mid X, W]=\theta(X)(T-h(X, W))+v=\theta(X) T_{\mathrm{res}}+v
\end{align*}
$$

@@

@kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield. However, when data are pooled across fields and applied nitrogen rates become continuous, the standard residualization of the treatment in DML does not extend cleanly to that setting.

@@

$$
\begin{align*}
Y=\theta(X, T)+g(X, W)+v
\end{align*}
$$

@@

$$
\begin{align*}
\theta(X, T)-\mathbb{E}[\theta(X, T) \mid X, W] \neq \theta(X, T-\mathbb{E}[T \mid X, W])
\end{align*}
$$

and the clean decomposition that underlies double-orthogonalization in the linear-in-$T$ case breaks down.

Instead, we apply single-orthogonalization (SO). We first estimate $f(X, W)=\mathbb{E}[Y \mid X, W]$ and form the residual $Y_{\mathrm{res}}=Y-f(X, W)$, and then recover $\theta(X, T)$ by flexibly modeling  $Y_{\mathrm{res}}$ as a function of $(X,T)$. This approach removes baseline variation due to $g(X,W)$ while preserving the full interaction structure between $X$ and the continuous treatment.

::: {.column-margin}
<span style = "color: blue;"> We will add justifications for this action later. </span>
:::
<!--end of column-margin-->

@@

$$
\begin{align*}
y = f(X, N) + g(W) + v
\end{align*}
$$

@@

$$
\begin{align*}
\tilde{y}=y-\widehat{g}(W)
\end{align*}
$$

@@

$$
\begin{align*}
\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
\end{align*}
$$
where $P_{\text {Corn }}$ is the corn price, $P_{\text {N }}$ is the nitrogen price, and $\hat{f}$ is the estimated yield response.

@@

::: {.column-margin}
<span style = "color: blue;"> Explain in  more detail about why it this is innovative. Explain the concept of linearization with splines using the figure I generated in my old manuscript. </span>
:::
<!--end of column-margin-->

@@

$$
\begin{align*}
y=\sum_{k=1}^K \theta_k(X) S_k(N)+g(X, W)+v
\end{align*}
$$

@@

$$
\begin{align*}
\tilde{y}=y-\mathbb{E}[y \mid X, W]=\sum_{k=1}^K \theta_k(X) \widetilde{S}_k(N)+v,
\end{align*}
$$

@@

$$
\begin{align*}
\tilde{y}=y-\widehat{\mathbb{E}}[y \mid X, W], \quad \tilde{N}=N-\widehat{\mathbb{E}}[N \mid X, W]
\end{align*}
$$

@@

$$
\begin{align*}
\hat{\tilde{y}}=\hat{\theta}(X) \cdot \tilde{N}
\end{align*}
$$

@@

$$
\begin{align*}
RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
\end{align*}
$$

@@

<span style = "color: red;"> Add profit differential as the measure. Discuss them in the results as well. </span>

@@

$$
\begin{align} 
y_{i,j} = 
  \begin{cases}
  \alpha_{i,j} + \beta_{i,j} N + \gamma_{i,j} N^2 + \varepsilon_{i,j}, & N < \tau_{i,j} \\
  \eta_{i,j} + \varepsilon_{i,j}, & N \geq \tau_{i,j}
  \end{cases}
\label{eq:yield-gen-sim}
\end{align}
$$

In this formulation, yield increases as $N_{i,j}$ increase until $N_{i,j}$ reaches $\tau_{i,j}$, at which yield hits the plateau, $\eta_{i,j}$. The yield level at $N_{i,j}= 0$ is $\alpha_{i,j}$. The rate at which yield increases with respect to $N_{i,j}$ is governed by $\beta_{i,j}$ and $\gamma_{i,j}$. For a given cell, the quadratic-plateu function has five parameters, but the degree of freedom is only three. That is, once any three of the five parameters are determined, then the rest of the parameters are automatically determined. Specifically, we first generated $\alpha_{i,j}$, $\eta_{i,j}$, and $\tau_{i,j}$ in a spatially correlated manner following the Gaussian bluh bluh process (cite), and then $\beta_{i,j}$ and $\gamma_{i,j}$ were derived. The erorr term ($\varepsilon_{i,t}$) is generated using the same spatial process. This is repeated independently 500 times to create 500 fields of the same size but different spatial distribution and values of the three parameters.

@@

```{r fig-trial-design, fig.cap = "Example trial design"}
g_td <- readRDS(here::here("writing/Figures/g_td.rds"))
```

@@

By contrast, NO_ANN exhibits substantial instability in the single-field case, with a heavy right tail and much higher errors, but rapidly converges toward the leading methods by 5–10 fields (mean RMSE improving to around 14–13), highlighting its strong dependence on sample size. SO_RF improves modestly with more fields (mid-20s RMSE for 10–20 fields) but remains consistently less accurate than SO_ANN or CF. NO_RF maintains broad, right-shifted distributions and higher mean errors throughout (≈ 36–40), confirming its role as a weak baseline. Finally, DO_ANN performs poorly throughout, with both histograms and means showing persistently high errors (≈ 90+) and negligible benefit from additional data, reflecting <span style = "color: red;"> over-correction </span>   in the orthogonalization process.

::: {.column-margin}
<span style = "color: blue;"> It is not over-correction. It is fundamentally messed up as the original functions are no longer retained. </span>
:::
<!--end of column-margin-->

@@

# ggsave(
#   filename = here::here("Proposal_presentation/fig_rmse_hist.png"),
#   plot = p,
#   width = 12,
#   height = 7,
#   bg = "white"
# )

@@

# ggsave(here::here("results","rmse_mean_by_model_fields.png"), p_mean, width = 9, height = 5, dpi = 300)

# ggsave(here::here("Proposal_presentation","model_improvement.png"), p_mean, width = 9, height = 5, dpi = 300)
```

@@

::: {.column-margin}
<span style = "color: blue;"> All this is probably redundant. What are you adding beyond the previous paragraphs? </span>
:::
<!--end of column-margin-->

@@

best_table <- best_models %>%
  gt() %>%
  tab_header(
    title = "Best models by field count",
    subtitle = "Ordered by lowest mean RMSE"
  ) %>%
  fmt_number(columns = c(mean), decimals = 3) %>%
  cols_label(
    field_f = "# fields",
    mean = "mean RMSE"
  ) %>%
  tab_options(table.font.size = "small")

#gtsave(best_table, here::here("Proposal_presentation/tbl_best_model_by_field_count.png"))

@@

<span style = "color: blue;"> I do not see the purpose of this much. Contrast CF, SO-NN, and the worst (all in the same panel) across the number of fields. </span>

### 2025-11-17 codes/R/Add_mean_match_to_YRF.qmd
comments on the manuscript added

>>>>

# CF

```{r}
# ==========================================================
# Mean-match predicted yield ONLY for CF (change # fields)
# ==========================================================
suppressPackageStartupMessages({
  library(here)
  library(readr)
  library(dplyr)
  library(stringr)
  library(purrr)
})

# Base directory
base_dir <- here::here("results", "yield_response_function_for_one_iteration")
target_dir <- file.path(base_dir, "YRF_CF_ten_fields")

if (!dir.exists(target_dir)) {
  stop("Target directory not found: ", target_dir, call. = FALSE)
}

# --- mean-match function (global shift per file) ---
mean_match_file <- function(csv_path) {
  df <- suppressMessages(readr::read_csv(csv_path, show_col_types = FALSE))

  if (!all(c("pred_yield", "true_yield") %in% names(df))) {
    message("Skipping (missing pred_yield/true_yield): ", csv_path)
    return(invisible(FALSE))
  }

  delta <- mean(df$true_yield, na.rm = TRUE) - mean(df$pred_yield, na.rm = TRUE)

# --- find CF yield_response_*.csv files and apply mean-match ---
csvs <- list.files(target_dir, pattern = "^yield_response_.*\\.csv$", full.names = TRUE)
if (!length(csvs)) {
  message("No yield_response_*.csv in: ", target_dir)
} else {
  walk(csvs, mean_match_file)
}

### 2025-11-17 codes/R/CF_YRF.qmd
comments on the manuscript added

# Post-estimation analysis
```{r}
cf_results <- readRDS(here::here("Results/yield_response_function_for_one_iteration/YRF_CF_10fields.rds"))
```

@@

```{r}
test_data <- cf_results$test_data
train_data <- cf_results$train_data
repsponse_data <- cf_results$repsponse_data

@@

  # pick a few ids
  check_id <- test_data[order(Nk), aunit_id] %>%
    .[seq(1, nrow(test_data), length = 5)] %>%
    .[-1]

@@

  # plot
  ggplot() +
    geom_line(
      data = curves_data_il[T <= 260, ],
      aes(y = yield, x = T, color = factor(group), linetype = type)
    ) +
    scale_color_discrete(name = "aunit_id Number") +
    scale_linetype_discrete(name = "Curve Type") +
    theme_bw() +
    theme(legend.position = "bottom") +
    ylab("Corn Yield (bu/acre)") +
    xlab("Nitrogen Application Rate (lb/acre)") +
    ggtitle(paste("Yield Response Function -", label))
}

file_labels <- c("1fields", "3fields", "5fields", "10fields", "20fields")
file_paths <- here::here(
  "Results/yield_response_function_for_one_iteration",
  paste0("YRF_CF_", file_labels, ".rds")
)

@@

# saveRDS(
#   all_plots,
#   here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF.rds")
# )

@@

all_plots <- readRDS(here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF.rds"))

>>>>

# Post-estimation analysis
```{r}
cf_results <- readRDS(here::here("Results/yield_response_function_for_one_iteration/YRF_CF_1fields.rds"))
```

@@

```{r}
test_data <- cf_results$test_data
train_data <- cf_results$train_data
response_data <- cf_results$response_data

@@

  # pick a few ids
  # check_id <- test_data[order(Nk), aunit_id] %>%
  #   .[seq(1, nrow(test_data), length = 2)] %>%
  #   .[-1]
   # Make CF use the same aunit_id as SO_ANN visualization
so_ann_csv <- here::here(
  "results", "yield_response_function_for_one_iteration",
  "YRF_SO_ANN_1_fields",   
  "yield_response_1.csv"
)

if (file.exists(so_ann_csv)) {
  so_df <- readr::read_csv(so_ann_csv, show_col_types = FALSE)
  # choose the same ID that SO_ANN used (the first unique aunit_id in its CSV)
  shared_id <- unique(so_df$aunit_id)[1]

  # ensure it exists in CF test_data
  if (shared_id %in% test_data$aunit_id) {
    check_id <- shared_id
  } else {
    # fallback: use the first available CF id
    check_id <- unique(test_data$aunit_id)[1]
  }
} else {
  # fallback if file missing
  check_id <- unique(test_data$aunit_id)[1]
}

message("Using aunit_id = ", check_id, " to match SO_ANN")

@@

  # plot

  ggplot() +
  geom_line(
    data = curves_data_il[T <= 260, ],
    aes(y = yield, x = T, color = factor(group), linetype = type)
  ) +
  scale_color_discrete(name = "aunit_id Number") +
  scale_linetype_discrete(name = "Curve Type") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylab("Corn Yield (bu/acre)") +
  xlab("Nitrogen Application Rate (lb/acre)") +
  ggtitle(paste("CF -", label))

# file_labels <- c("1fields", "3fields", "5fields", "10fields", "20fields")
# file_paths <- here::here(
#   "results/yield_response_function_for_one_iteration",
#   paste0("YRF_CF_", file_labels, ".rds")
# )

field_counts <- c(1, 3, 5, 10, 20)

file_paths <- here::here(
  "results/yield_response_function_for_one_iteration",
  sprintf("YRF_CF_%dfields.rds", field_counts)
)

file_labels <- sprintf("%d fields", field_counts)

@@

# saveRDS(
#   all_plots,
#   here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF_double.rds")
# )

@@

all_plots <- readRDS(here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF_double.rds"))

### 2025-11-17 codes/R/GRF_CF_Loop.qmd
comments on the manuscript added

---
title: "CF Loop"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

@@

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
)

#--- python functions ---#
use_condaenv("cf_conda311", required = TRUE)

# use_condaenv("cf_conda311",
#              conda = "/Users/monamousavi/miniforge3/bin/conda",
#              required = TRUE)

py_config()
py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")

@@

#--- python functions ---#
source_python(here::here("codes/Python/run_CF_c.py"))
```

## load data

```{r}
#raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))

# how many fields to train on?
n_fields <- 20

splits_file <- here(
  "data", "train_test_split",
  sprintf("train_test_splits_%dfields.csv", n_fields)
)
output_dir <- here(
  "results",
  sprintf("CF_outcome_%d_fields", n_fields)
)
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

@@

# ─── LOAD SPLITS ────────────────────────────────────────────────────────────────

train_test_splits <- read_csv(splits_file)

@@

# CF loop

```{r}
# ── EONR subfolder ────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

@@

>>>>

---
title: "GRF CF Loop"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

@@

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1" 
)

@@

## load data
```{r}
#raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))

# how many fields to train on?
n_fields <- 5

splits_file <- here(
  "data", "train_test_split",
  sprintf("train_test_splits_%dfields.csv", n_fields)
)
output_dir <- here(
  "results/GRF_CF",
  sprintf("CF_outcome_%d_fields", n_fields)
)
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

@@

# ─── LOAD SPLITS ────────────────────────────────────────────────────────────────
train_test_splits <- read_csv(splits_file)

@@

# CF loop
```{r}
# ── EONR subfolder ────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

@@

```{r}

# ── EONR subfolder ──────────────────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

## ── Loop over test sims ───────────────────────────────────────────────────────
#for (i in seq_len(1))
#for (i in seq_len(nrow(train_test_splits)))
for (i in seq_len(1)) {
  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  tryCatch({
    ## 1) train/test IDs
    split_row     <- train_test_splits[i, ]
    train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
    sim_id_ls     <- c(train_sim_ids, test_sim_id)

    ## 3) CF variables
    x_vars <- c("Nk", "plateau", "b0")
    T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
    Y      <- train_data[, yield]
    X      <- as.matrix(train_data[, ..x_vars])
    W      <- X
    X_test <- as.matrix(test_data[, ..x_vars])

    ## 4) cf in GRF
    cf_fit <- causal_forest(
      X = X,
      Y = Y,
      W = T_info$T_sp,     # spline
      num.trees = 2000,
      min.node.size = 10,
      seed = 3747823
    )

    # Estimate treatment effects for test data
    te_hat_cf <- predict(cf_fit, newdata = X_test)$predictions

    te_info <- data.table(aunit_id = test_data$aunit_id, est = te_hat_cf)

    csv_path <- file.path(eonr_dir, "eonr_progress.csv")
    data.table::fwrite(e, csv_path, append = file.exists(csv_path))

    ## 8) Save results object (optional)
    # res <- list(
    #   cf_fit        = cf_fit,
    #   te_info       = te_info,
    #   T_seq         = T_seq,
    #   response_data = response_data,
    #   test_data     = test_data,
    #   train_data    = train_data,
    #   combined_eonr = combined_eonr,
    #   corr          = corr_val
    # )
    # tmp_fn <- paste0(out_fn, ".tmp")
    # saveRDS(res, tmp_fn, compress = "xz")
    # file.rename(tmp_fn, out_fn)

    message("Finished sim ", test_sim_id,
            " | corr = ", round(corr_val, 3))

    # Free RAM
    rm(cf_fit, te_info, response_data, train_data, test_data, e)
    if (i %% 10 == 0) gc()
  },
  error = function(e) {
    message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
    err_row <- data.frame(
      test_sim_id = test_sim_id,
      error       = conditionMessage(e),
      ts          = as.character(Sys.time())
    )
    write.table(
      err_row,
      file      = file.path(output_dir, "errors.csv"),
      sep       = ",",
      row.names = FALSE,
      col.names = !file.exists(file.path(output_dir, "errors.csv")),
      append    = TRUE
    )
  })
}
